systemfit: A Package to Estimate\\ Simultaneous Equation Systems in \proglang{R}

### 2006-02-18 paper/trunk/nlsystemfit.tex
moved subsection about non-linear estimation from section "other issues" to the file "nlsystemfit.tex" that is (currently) not included in the article; added paragraph about nlsystemfit in section "summary and outlook"; removed "non-linear" from section "introduction"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@612 591db2bd-e0a0-4d22-8c19-2192dbb7da31

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Nonlinear estimation}

As linear systems of equations can be estimated using
\code{systemfit}, nonlinear systems of equations \citep{hausman75}
can be estimated using a similar function; \code{nlsystemfit}.
The usage for the nonlinear estimation function is similar to
\code{systemfit} and can estimate system of equations using OLS, SUR,
2SLS and 3SLS methods.
The equations are defined as,

\begin{equation}
  \label{eq:non_linear_eq_1}
  \epsilon_{t} = q( y_t, x_t, \beta )
\end{equation}

\begin{equation}
  \label{eq:non_linear_eq_2}
  z_{t} = Z( x_t )
\end{equation}
where $\epsilon_{t}$ are the residuals from the $y_t$ observations,
$x_t$ is a matrix of independent variables, $\beta$ is a vector is
parameter estimates and $z_{t}$ are the functions evaluated at the
parameter estimates.

where \code{method} is one of the following estimation methods: "OLS",
"SUR", "2SLS", or "3SLS", \code{eqns} is a list of equations similar
to those described for \code{systemfit}, and \code{start} is a list of
starting values of the parameter estimates.
The nlsystemfit function relies on \code{nlm} to perform the minimization
of the objective functions and the \code{qr} set of functions.
If the user does not have a set of estimates for the initial parameters,
it is suggested using linearized forms in one of the linear methods,
or simply using \code{nls} to obtain estimates for the equations.
The outputs are similar to those obtained by \code{systemfit} objects.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% left over from the older docs
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% The mandatory argument
% \code{eqns} is a list of the equations to estimate.  Each equation is
% a standard formula in \proglang{R}.  It starts with a dependent
% variable on the left hand side.  After a tilde ($~$) the regressors
% are listed  \footnote{For Details see the \proglang{R} help files to
%   \code{formula}}.

% The objective functions for the methods are:

% \begin{center}
% \begin{tabular}{|l|c|c|c|} \hline
%   Method & Instruments & Objective Function & Covariance of $\theta$ \\ \hline  
%   OLS & No & $r'r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline
%   SUR & No & $r'(diag(S)_{OLS}^{-1}\otimes I)r$ & $(X(S^{-1}\otimes I)X)^{-1}$ \\ \hline
%   2SLS & Yes & $r'(I \otimes W)r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline 
%   3SLS & Yes & $r'(S_{2SLS}^{-1} \otimes W)r$ & $(X(diag(S)^{-1}\otimes W)X)^{-1}$ \\ \hline
% \end{tabular}
% \end{center}

% where, $r$ is a column vector for the residuals for each equation,
% what is the difference to \epsilon?
% $X$ is matrix of the partial derivatives of the dependent variable 
% with respect to the parameters $\left( \frac{ \partial y }{ \theta} \right)$,
% is this correct?
% in the linear section X is used for the regressors, 
% thus we should use a different variable name here.
% $W$ is a matrix of the instrument variables, $Z(Z'Z)^{-1}Z$, $Z$ is a
% matrix of the instrument variables, and $I$ is an $n \times n $
% identity matrix and $S$ is the estimated covariance matrix between the
% equations

% \begin{equation}
%   \label{eq:non-linear_varcov}
%   \hat{\sigma}_{ij} = (\hat{e}_i' \hat{e}_j) / \sqrt{(T - k_i)*(T- k_j)} 
% \end{equation}

% Do we actually have the covariance function in place.

% The argument \code{method} is a string determining the estimation method.
% It must be one of "OLS", "WLS", "SUR", "2SLS", "W2SLS" or "3SLS".

% \textbf{The residual covariance matrix can be calculated in
%   different ways (section~\ref{sec:residcov}).  It should be
%   relatively easy to implement this also in nlsystemfit(),  e.g. the
%   function 'calcRCov' in systemfit() could be moved outside
%   systemfit() that it can be used also by nlsystemfit().}

% \textbf{You need to clear this up.}

### 2006-02-18 paper/trunk/systemfit_intro.tex
moved subsection about non-linear estimation from section "other issues" to the file "nlsystemfit.tex" that is (currently) not included in the article; added paragraph about nlsystemfit in section "summary and outlook"; removed "non-linear" from section "introduction"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@612 591db2bd-e0a0-4d22-8c19-2192dbb7da31

The \pkg{systemfit} package provides the capability to estimate
linear and non-linear equation systems in \proglang{R}
\citep{r-project}.
Although linear and non-linear equation systems can be estimated
with several other statistical and econometric software packages
(e.g.\ \proglang{SAS}, \proglang{EViews}, \proglang{TSP}),
\pkg{systemfit} has several advantages.
First, all estimation procedures are publicly available in the source code.
Second, the estimation algorithms can be easily modified to meet specific
requirements.
Third, the (advanced) user can control estimation details generally
not available in other software packages by overriding reasonable defaults.

>>>>

The \pkg{systemfit} package provides the capability to estimate
linear equation systems in \proglang{R}
\citep{r-project}.
Although linear equation systems can be estimated
with several other statistical and econometric software packages
(e.g.\ \proglang{SAS}, \proglang{EViews}, \proglang{TSP}),
\pkg{systemfit} has several advantages.
First, all estimation procedures are publicly available in the source code.
Second, the estimation algorithms can be easily modified to meet specific
requirements.
Third, the (advanced) user can control estimation details generally
not available in other software packages by overriding reasonable defaults.

### 2006-02-18 paper/trunk/systemfit_other.tex
added section about how to use systemfit to test linear restrictions; changed restriction: beta_2 = -beta_5 -> beta_2 = -beta_6; a very few other corrections
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@607 591db2bd-e0a0-4d22-8c19-2192dbb7da31
moved subsection about non-linear estimation from section "other issues" to the file "nlsystemfit.tex" that is (currently) not included in the article; added paragraph about nlsystemfit in section "summary and outlook"; removed "non-linear" from section "introduction"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@612 591d

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Testing Linear Restrictions}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Nonlinear estimation}

As linear systems of equations can be estimated using
\code{systemfit}, nonlinear systems of equations \citep{hausman75}
can be estimated using a similar function; \code{nlsystemfit}.
The usage for the nonlinear estimation function is similar to
\code{systemfit} and can estimate system of equations using OLS, SUR,
2SLS and 3SLS methods.
The equations are defined as,

\begin{equation}
  \label{eq:non_linear_eq_1}
  \epsilon_{t} = q( y_t, x_t, \beta )
\end{equation}

\noindent and

\begin{equation}
  \label{eq:non_linear_eq_2}
  z_{t} = Z( x_t )
\end{equation}
where $\epsilon_{t}$ are the residuals from the $y_t$ observations,
$x_t$ is a matrix of independent variables, $\beta$ is a vector is
parameter estimates and $z_{t}$ are the functions evaluated at the
parameter estimates.

Similar to calling the \code{systemfit} function, \code{nlsystemfit}
is called with a minimum of three arguments,

\code{
R> nlsystemfit( method, eqns, start )
}

where \code{method} is one of the following estimation methods: "OLS",
"SUR", "2SLS", or "3SLS", \code{eqns} is a list of equations similar
to those described for \code{systemfit}, and \code{start} is a list of
starting values of the parameter estimates.
The nlsystemfit function relies on \code{nlm} to perform the minimization
of the objective functions and the \code{qr} set of functions.
If the user does not have a set of estimates for the initial parameters,
it is suggested using linearized forms in one of the linear methods,
or simply using \code{nls} to obtain estimates for the equations.
The outputs are similar to those obtained by \code{systemfit} objects.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% left over from the older docs
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% The mandatory argument
% \code{eqns} is a list of the equations to estimate.  Each equation is
% a standard formula in \proglang{R}.  It starts with a dependent
% variable on the left hand side.  After a tilde ($~$) the regressors
% are listed  \footnote{For Details see the \proglang{R} help files to
%   \code{formula}}.

% The objective functions for the methods are:

% \begin{center}
% \begin{tabular}{|l|c|c|c|} \hline
%   Method & Instruments & Objective Function & Covariance of $\theta$ \\ \hline  
%   OLS & No & $r'r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline
%   SUR & No & $r'(diag(S)_{OLS}^{-1}\otimes I)r$ & $(X(S^{-1}\otimes I)X)^{-1}$ \\ \hline
%   2SLS & Yes & $r'(I \otimes W)r$ & $(X(diag(S)^{-1}\otimes I)X)^{-1}$ \\ \hline 
%   3SLS & Yes & $r'(S_{2SLS}^{-1} \otimes W)r$ & $(X(diag(S)^{-1}\otimes W)X)^{-1}$ \\ \hline
% \end{tabular}
% \end{center}

% where, $r$ is a column vector for the residuals for each equation,
% what is the difference to \epsilon?
% $X$ is matrix of the partial derivatives of the dependent variable 
% with respect to the parameters $\left( \frac{ \partial y }{ \theta} \right)$,
% is this correct?
% in the linear section X is used for the regressors, 
% thus we should use a different variable name here.
% $W$ is a matrix of the instrument variables, $Z(Z'Z)^{-1}Z$, $Z$ is a
% matrix of the instrument variables, and $I$ is an $n \times n $
% identity matrix and $S$ is the estimated covariance matrix between the
% equations

% \begin{equation}
%   \label{eq:non-linear_varcov}
%   \hat{\sigma}_{ij} = (\hat{e}_i' \hat{e}_j) / \sqrt{(T - k_i)*(T- k_j)} 
% \end{equation}

% Do we actually have the covariance function in place.

% The argument \code{method} is a string determining the estimation method.
% It must be one of "OLS", "WLS", "SUR", "2SLS", "W2SLS" or "3SLS".

% \textbf{The residual covariance matrix can be calculated in
%   different ways (section~\ref{sec:residcov}).  It should be
%   relatively easy to implement this also in nlsystemfit(),  e.g. the
%   function 'calcRCov' in systemfit() could be moved outside
%   systemfit() that it can be used also by nlsystemfit().}

% \textbf{You need to clear this up.}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Testing Linear Restrictions}
\label{sec:testingRestrictions}

@@

### 2006-02-18 paper/trunk/systemfit_summary.tex
moved subsection about non-linear estimation from section "other issues" to the file "nlsystemfit.tex" that is (currently) not included in the article; added paragraph about nlsystemfit in section "summary and outlook"; removed "non-linear" from section "introduction"
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/systemfit@612 591db2bd-e0a0-4d22-8c19-2192dbb7da31

>>>>

\paragraph{Non-linear estimation}
The \pkg{systemfit} package provides a function to estimate
systems of non-linear estimations.
However, the function \code{nlsystemfit} is currently under development
and the results are not yet always reliable
due to convergence problems.
We hope to improve this behavior soon.
